---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# AustralianElections

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/RohanAlexander/AustralianElections.svg?branch=master)](https://travis-ci.org/RohanAlexander/AustralianElections)
<!-- badges: end -->

`AustralianElections` is a collection of datasets related to Australian elections The datasets are:

* **elections.rda**: The main dataset.
* **parliaments.rda**: Dates of each parliament.
* **voting_data.rda**: All federal elections. 

The datasets are up-to-date as of 10 September 2019.

If you have suggestions on how I could improve the datasets, or corrections, please don't hesitate to get in touch.


## Installation

You can install this package from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("RohanAlexander/AustralianElections")
```

## Example

This is a example of how to load the data:

```{r loaddata, eval=FALSE, echo=TRUE}
library(tidyverse)

devtools::install_github("RohanAlexander/AustralianElections")

elections <- AustralianElections::elections %>% as_tibble()
parliaments <- AustralianElections::parliaments %>% as_tibble()
voting_data <- AustralianElections::voting_data %>% as_tibble()
```


## Dataset details

TBD

## TODO

## Roadmap



## Sources


## Acknowledgements


## Citation

If you use `AustralianElections`, please consider citing:

Alexander, Rohan. (2019). AustralianElections: Datasets on Australian Elections. Source: https://github.com/RohanAlexander/AustralianElections.

## Author information

**Rohan Alexander** (corresponding author and repository maintainer)  
University of Toronto  
Information Sciences  
140 St George St  
Toronto, ON, Canada  
Email: rohan.alexander@utoronto.ca

